Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants

### 2018-06-21 paper/reviewer_response_PNAS.tex

4) Were sequences from passaged viruses (isolated on MDCKs and other cell lines) included in the comparison between in vitro mutagenesis data and natural evolution data? They need to be excluded since their presence might induce a bias towards what is tolerated/preferred in cell culture. Only sequences obtained directly from patient specimens should be used for the comparison.

\response{We thank the reviewer for this astute suggestion and now include an analysis of the relationship between mutational effect and maximum frequency for mutations from unpassaged isolates.
Due to the sparseness of unpassaged isolate sequences previous to the last decade, we analyzed unpassaged isolates from the post-Perth/2009 partition of the H3N2 HA phylogenetic tree.
Figures 5, 6, and S7 now include panels showing the results for these unpassaged isolate mutations.
The results in Figure 5 demonstrate that the correlation between mutational effect as determined by our experiments and maximum mutation frequency is stronger for unpassaged isolates than when we include all isolates ($\rho = 0.24$).
We do not see such a correlation when we use the WSN/1933 H1 HA preferences to estimate the mutational effects for unpassaged isolate mutations.
}

5) It is unclear how the ``rescaling'' is performed and what it exactly does. The authors need to described and justify this better and should also provide the non-rescaled data in a ``sequence logo'' figure for comparison.

\response{Re-scaling of the preferences is performed by fitting a stringency parameter that is a reflection of how strongly our experimentally determined preferences correspond to the gene's evolution in nature.
This parameter is fit under an experimentally informed codon substitution model (ExpCM) using the phydms program (Hilton 2017 PeerJ).
A stringency parameter greater than 1 indicates that natural selection favors the same amino acids as the experiments but with greater stringency, while a value less than 1 indicates that natural selection does not favor the amino acids with as much stringency as the experiments.
Because selection in the lab often does not fully reflect the stringency of selection in nature and selection across experiments may not be entirely consistent, we re-scale the preferences by this stringency parameter.
Importantly, re-scaling does not affect any of the evolutionary analyses performed in this paper since it is a constant that multiplies all mutational effects, simply corresponding a change in units on the x-axis of Figure~5.
We now include a supplementary panel A to Figure S3 showing the un-scaled Perth/2009 H3 HA preferences.
}

@@

d) The supernatant is never centrifuged before RNA extraction. Cell debris is not removed which is likely contributing significantly to RNA sequences detected in the supernatant. However, RNA associated with cell debris is not representative of RNA from fit, replicating viruses.

\response{We clarified the viral supernatant before RNA extraction by centrifuging the supernatant at 2000 $\times$ \textit{g} for five minutes to remove any cellular debris.
We failed to mention this step in the Methods, and again deeply apologize for this critical oversight.
We have updated the Methods to now include mention of this step.
}

The experiments should be repeated with at least one more passaging step, proper removal of inoculum and washing of the cells after infection. Centrifugation steps should also be included between the passages and before RNA extraction in the end.

\response{For the reasons described above, it is not necessary to totally repeat the experiments.
Certainly other experimental parameter choices could have been used, but this is always the case for complex experiments.
Our choices are reasonable, and the four lines of evidence listed at the beginning of this comment show that they yield reproducible and meaningful results.}

@@

1) Many abbreviations are not defined in the manuscript including MDCK-SIAT1, TMPRSS2, MOI, DMEM, FBS, LB.

\response{We now explain these abbreviations in the manuscript.}

2) The authors state on page 5 that mutations detected in cell culture as deleterious rarely reach high frequency in nature. Could you please give a few examples?

\response{We individually validated viral growth of a C52A mutant, which removes a Cys critical in forming a disulfide bond with Cys-277 and maintaining HA structure.
We confirmed that this mutant grows very poorly when rescued, and mutations at site 52 never arise at a detectable level.

As another example, mutation N170H fails to reach $>$1\% frequency in nature, and according to our dataset, is a mutation towards a highly unpreferred amino acid. 
}

@@

\comment{Not obvious how to give an example, as there are many mutations we identify as deleterious and many mutations that don't reach high frequency. 
But I suppose if we validate some specific mutation (e.g., a cysteine mutation) as deleterious, we can point out that it doesn't reach high frequency...}

3) The authors picked 31 random clones to evaluate the mutation rate per clone. These sequences should be made available. Did this set include viruses without a proper start codon (see above)? What was their growth phenotype in vitro?

\response{We have now added the list of mutations from the 31 clones to Dataset S3.
Two of the clones had a mutation that removed the ATG start codon.
One of these resulted in an M(-16)V mutation, while the other resulted in an M(-16)Q mutation which was linked to a Q80I mutation.
According to our preferences, Val and Gln are not very preferred at this site, suggesting that these mutants would not grow well compared to wildtype.
}

@@

\subsubsection*{Major points:}

The authors themselves acknowledge the most important limitation of this study: that the experimental measure of viral growth in cell culture is not reflective of true fitness in nature. Most importantly, the work presented here does not capture host-induced pressures on the antigenic regions of the viral protein. This is evidenced by the discrepancy between the high mutational tolerance they calculated for the stalk region and the actual change in stalk amino acid sequence. A simple protein BLAST comparison between the amino acid sequences of HA from H3N2 viruses Perth09 and HK68 shows a 77\% conservation in head amino acids but a 94\% conservation in stalk amino acids. Additionally, according to figure 2, there is no observable reversion of the G78D mutation they introduced despite the dominance of G78 in the population as seen in figure S1. The authors recognize this in their work and do not attempt to exaggerate the significance of these findings. Still, it would be useful to differentiate between head and stalk mutations when depicting individual amino acid mutations in figure 4 and their frequency trajectories to highlight that the predominance of mutations that occur in nature are in the head.

\response{We thank the reviewer for acknowledging our efforts to fairly present the considerable caveats associated with cell-culture measurements.

The suggestion to make a version of Fig.~4 that distinguished between the stalk and the head is a good one.
It is difficult to also overlay the structural location (stalk versus head) on the current Fig.~4, since colors are already used to indicate mutational effects.
We have therefore created a new supplemental figure (Fig.~S8) that shows these data---as the reviewer suspected, this shows that head mutations are much more common.

The reviewer is correct that G78D appears to be a site where the correspondence between selection in our experiments and frequency in nature is low.
This makes sense, as G78D is one of the two sites that we specifically identified as being a lab adaptation (see Figure~S1).
We now explicitly mention this fact in the Discussion as follows:

\begin{quote}
\textsl{Indeed, a vast amount of work in virology has chronicled the ways in which experiments can select for lab artifacts or fail to capture important pressures that are relevant in nature.
As an example of this, although we identified G78D as a lab adaptation mutation and the wildtype Gly is not highly preferred at this site, this mutation never reaches above 20\% frequency in nature. }
\end{quote}
}

@@

The data regarding greater mutational tolerance in the stalk compared to the head should be paired with data from deep mutational scanning utilizing stalk- and head-specific antibodies as a selection process as this group has previously done with WSN H1 (Doud et al., Nat Comm 2018, Doud et al., PLoS Path 2017) to see if tolerant sites are responsible for viral escape.

\response{We agree that this would be a very interesting experiment to perform.
However, we believe such a study, which would require rigorous examination of escape from multiple head- and stalk-specific antibodies, is beyond the scope of this manuscript.
}

Additionally, the authors should directly show the discrepancies between this in vitro selection and the H3 HA's evolution in nature. This group has performed such an analysis in their previous deep mutational scanning work with the H1 HA of WSN (Doud and Bloom, Viruses 2016). Maximum likelihood phylogenetics can be used to estimate the difference between experimental amino acid preference and natural preference. One might expect that this would provide some insight into the nature of the bias introduced through cell culture.

\response{This is a good suggestion and we have now added the results of this analysis in Figure S4.
As the logo plot shows, although there are some differences, at most sites there is little difference between the amino acids that are preferred in our experiments to their frequencies in nature.
}

@@

The authors are upfront about the limited degree to which their calculated mutational effect and the maximum mutation frequency in nature correlate, describing it as ``modest.'' One of the major points of the paper is that conclusions from this type of analysis cannot be generalized when using data from strains that are too divergent. The authors should address the possibility that data from deep mutational scanning can be generalized to divergent strains if the study is done in a system that is more reflective of nature.

\response{We now have acknowledged this possibility in the Discussion in the sentence ``We suspect that performing similar experiments using more realistic and complex selections (e.g., ferrets or primary human airway cultures) might further improve their utility and possibly their generalizability to more divergent strains."}

@@

The experimental approach of deep mutational scanning has been applied to viral proteins before, including HIV env, influenza H1 HA, and NP. The novelty in this paper is that it is the first time such an analysis has been performed for H3 specifically and that it demonstrates the (albeit limited) utility of deep mutational scanning in predicting the success of viral strains in nature. Ultimately, it is clear from the discussion that the authors see this study as an intermediate step, suggesting that their work is "an improvement over no information at all." While something is indeed better than nothing, in my opinion, this paper represents only an incremental step forward in the field of influenza evolutionary forecasting due to the aforementioned limitations, especially since this experiment and subsequent analysis could be performed using more stringent selection methods during viral passaging like treatment with human sera for data that is presumably more faithful to nature.

\response{This is a fair assessment by the reviewer.
However, this is the first paper to show that data from deep mutational scanning can actually reflect evolutionary outcomes in nature, and provides some insight into the extent by which functional constraints on HA shape its evolution. 
Furthermore, it is possible that this work can provide a useful framework for applying deep mutational scanning studies in the context of evolution in nature.
}

@@

\subsubsection*{Minor Points:}

The correlation for prospective analysis (post-Perth) is smaller than the correlation for post-hoc analysis (pre-Perth). Since prospective analysis is more important when predicting evolutionary fates, it might be beneficial for the authors to comment on this difference.

\response{We now include an analysis of mutational effects and maximum frequency for mutations of unpassaged isolates (see Figures 5, 6, and S7) from the post-Perth partition of the phylogenetic tree to limit any potential biases that our experimental measurements may have towards mutations from cell- or egg-passaged isolates.
The results show that this correlation is stronger when considering mutations from unpassaged isolates in a post-hoc analysis, demonstrating the potential utility of our experimental measurements in a prospective analysis, specifically when passaged isolate sequences are excluded.
}

@@

A number of issues interfere with the clarity of this manuscript, particularly when appealing to a broader scientific audience. While this is understandable given the nature of this study, several adjustments should be made. A clear description of mutational effect is never provided in the results, though it forms a crucial component of the data. As such, it is not really intuitive as to why the range from -10-0 should be considered "neutral." Additionally, mutational tolerance is quantified as Shannon entropy of the re-scaled amino-acid preferences at each site, but it is unclear how this entropy is calculated. The entire first half of the second results section describing how the experimentally informed codon substitution model they generate is better than existing models and the associated Table 1 is very difficult to understand. AIC is never explained. The significance of dN/dS being much less than 1 for conventional models should be explained. The purpose of modeling of ExpCM, site avg. is unclear.

\response{We appreciate these suggestions from the reviewer and have made the following changes:
\begin{itemize}
\item We have now included a description of how we calculated mutational effects in the Results section.
\item We agree that a cutoff of -10 to 0 is arbitrary and removed the sentence ``Mutations that reach high frequencies generally have neutral or beneficial effects according to our experimental measurements." 
\item We included in the Methods a description of how we calculated Shannon entropy.
\item We now define AIC in the legend of Table 1.
\item We have now included an explanation of dN/dS $\ll$1 for conventional substitution models, and now describe the utility of building an ExpCM, site avg. substitution model
\end{itemize}
}

@@

In figure 4, it would be useful to show Perth09 on the tree and mention that closely related strains are excluded from analysis in the figure legend. As it stands, there is just an unlabeled cluster of viruses for which no individual amino acid mutations are characterized. It is also important for the authors to clearly state what frequency they are talking about. Specifically, in consideration of a particular mutation, what is the population in the denominator?

\response{We have now labeled the Perth/2009 strain on the tree in Figure 4 and now mention in the caption that we have excluded closely related strains from the analysis...
}

\comment{Fixed the figure. Maybe John and Trevor can do a better job than me at explaining the frequency denominator. Is it the population observed at a given pivot?}

### 2018-06-21 paper/manuscript.tex
Finish drafting Reviewer Responses, update manuscript revisions

\begin{table}[b!]
\caption{\label{tab:phydms}
{\bf Substitution models informed by the experiments describe HA's evolution better than traditional models.}}
\begin{center}
\begin{tabular}{cccccccc}
\hline
\bf{Model} & \bf{$\Delta$AIC} & \bf{LnL} & \bf{Stringency} & \bf{$\omega$}  \\ \hline
ExpCM & 0.0 & -8441 & 2.47 & 0.91 \\
GY94 M5 & 2094 & -9482 & -- & 0.36 (0.30, 0.84) \\
ExpCM, site avg. & 2501 & -9692 & 0.67 & 0.32 \\
GY94 M0 & 2536 & -9704 & -- & 0.31 \\
\hline
\end{tabular}
 \end{center}
\addtabletext{Maximum likelihood phylogenetic fit to an alignment of human H3N2 HAs using ExpCM~\cite{hilton2017phydms}, ExpCM in which the experimental measurements are averaged across sites (site avg.), and M0 and M5 versions of the Goldman-Yang (GY94) model~\cite{yang2000codon}.
Models are compared by AIC~\cite{posada2004model} computed from the log likelihood (LnL) and number of model parameters.
The $\omega$ parameter is dN/dS for the Goldman-Yang models, and the relative dN/dS after accounting for the measurements for the ExpCM.
For the M5 model, we give the mean followed by the shape and rate parameters of the gamma distribution over $\omega$.
}
\end{table}

@@

\subsection*{Our measurements are consistent with existing knowledge about HA's evolution and function}
How do the HA amino-acid preferences measured in our experiments relate to the evolution of H3N2 influenza virus in nature?
This question can be addressed by evaluating how well an experimentally informed codon substitution model (ExpCM) using our measurements describes H3N2 evolution compared to standard substitution models~\cite{bloom2017identification,hilton2017phydms}.
Table~\ref{tab:phydms} shows that the ExpCM using the across-replicate average of our measurements greatly outperforms conventional substitution models.
This result indicates that our experiments authentically capture some of the constraints on HA evolution.
The relative rate of nonsynonymous to synonymous substitutions (dN/dS or $\omega$) is $\ll$1 for conventional substitution models (Table~\ref{tab:phydms}).
However, the relative rate of nonsynonymous to synonymous substitutions after accounting for the amino-acid preferences measured in our experiments ($\omega$ for the ExpCM) is close to one (Table~\ref{tab:phydms}), indicating that most purifying selection against nonsynonymous substitutions is accounted for in the deep mutational scanning.
The ExpCM stringency parameter~\cite{hilton2017phydms} is 2.47 (Table~\ref{tab:phydms}), indicating that natural selection favors the same amino acids as the experiments but with greater stringency.
Throughout the rest of this paper, we use the amino-acid preferences re-scaled~\cite{bloom2017identification,hilton2017phydms} by this stringency parameter.
These re-scaled preferences are shown in Figure~\ref{fig:logoplot}.

@@

After annotating mutations and their frequencies on the phylogeny in this way (Figure~\ref{fig:frequency_trajectory}), it is visually obvious that there are relatively few circulating mutations that we measure to be strongly deleterious---and that such deleterious mutations rarely reach high frequency when they do occur.
Mutations that reach high frequencies generally have neutral or beneficial effects according to our experimental measurements.

We next sought to quantify the correlation between a mutation's experimentally measured effect and the maximum frequency it attained during natural evolution.
To minimize effects related to the genetic background of the strain used in the experiment, we excluded mutations closely related to the experimental strain itself and partitioned the remaining mutations into 1,022 mutations pre-dating and 299 mutations post-dating the Perth/2009 strain (Figure~\ref{suppfig:tree}).
We additionally excluded mutations from the post-Perth partition that were sampled in 2014 or after that have not had enough time for their evolutionary fates to be fully resolved.
We used these pre-Perth and post-Perth partitions to test the utility of our measurements for both post-hoc and prospective analyses, respectively.
We quantified the relationship between mutational effects and maximum mutation frequencies in the H3N2 phylogeny via Spearman rank correlation (Figure~\ref{fig:muteffect_maxfreq}A).
In both pre-Perth and post-Perth time periods, we found a modest, but statistically significant relationship between mutational effect and maximum mutation frequency (pre-Perth $\rho = 0.17$, post-Perth $\rho = 0.14$). 
To restrict potential biases of our experimental measurements towards mutations from cell- or egg-passaged isolates, we evaluated mutations from unpassaged isolates.
Due to the limited availability of primary isolate sequences until the last decade, we analyzed mutations of unpassaged isolates from the post-Perth partition of the phylogenetic tree~\cite{mcwhite2016sequence}.
The correlation between mutational effect and maximum frequency was stronger for mutations from unpassaged post-Perth isolates ($\rho = 0.24$).
The similar effect sizes for both partitions shows that our experimental measurements can help explain the evolutionary fates of mutations in strains that post-date the experimentally studied strain, as well as to retrospectively analyze mutations that precede the experimental strain.

@@

It is important to emphasize that measurements of viral growth in cell culture do \emph{not} represent true fitness in nature.
Indeed, a vast amount of work in virology has chronicled the ways in which experiments can select for lab artifacts or fail to capture important pressures that are relevant in nature~\cite{daniels1985fusion,sun2010modifications,lee2013comparison,wu2017structural}.
Mutations in viral genes other than HA are also important in determining strain success~\cite{memoli2009recent,raghwani2017selection}.
Given these caveats, it might seem surprising that measuring viral growth in cell culture can be informative about the success of viral strains in nature.
Yet, prior to our work, there were no comprehensive studies of the functional effects of mutations to H3 HA on any property that even resembled viral fitness in nature, and modeling work has either omitted the non-antigenic effects of mutations~\cite{sun2013using,harvey2016identification,neher2016prediction} or assumed that all non-epitope mutations had equivalent deleterious effects~\cite{luksza2014predictive}.
The strength of our measurements are not that they perfectly capture fitness in nature, but that they are systematic and quantitative---and so represent an improvement over no information at all.
We suspect that performing similar experiments using more realistic and complex selections (e.g., ferrets or primary human airway cultures) might further improve their utility and possibly their generalizability to more divergent strains.

@@

\subsection*{Quantification of relative solvent accessibility}
We quantified the absolute solvent accessibility of each site of the H3 HA (PDB 4O5N;~\cite{lee2014receptor}) or the H1 HA (PDB 1RVX;~\cite{gamblin2004structure}) structure using DSSP~\cite{kabsch1983}.
We then normalized to a relative solvent accessibility using the absolute accessibilities in~\cite{tien2013}.

>>>>

\begin{table}[b!]
\caption{\label{tab:phydms}
{\bf Substitution models informed by the experiments describe HA's evolution better than traditional models.}}
\begin{center}
\begin{tabular}{cccccccc}
\hline
\bf{Model} & \bf{$\Delta$AIC} & \bf{LnL} & \bf{Stringency} & \bf{$\omega$}  \\ \hline
ExpCM & 0.0 & -8441 & 2.47 & 0.91 \\
GY94 M5 & 2094 & -9482 & -- & 0.36 (0.30, 0.84) \\
ExpCM, site avg. & 2501 & -9692 & 0.67 & 0.32 \\
GY94 M0 & 2536 & -9704 & -- & 0.31 \\
\hline
\end{tabular}
 \end{center}
\addtabletext{Maximum likelihood phylogenetic fit to an alignment of human H3N2 HAs using ExpCM~\cite{hilton2017phydms}, ExpCM in which the experimental measurements are averaged across sites (site avg.), and M0 and M5 versions of the Goldman-Yang (GY94) model~\cite{yang2000codon}.
Models are compared by Akaike information criterion (AIC)~\cite{posada2004model} computed from the log likelihood (LnL) and number of model parameters.
The $\omega$ parameter is dN/dS for the Goldman-Yang models, and the relative dN/dS after accounting for the measurements for the ExpCM.
For the M5 model, we give the mean followed by the shape and rate parameters of the gamma distribution over $\omega$.
}
\end{table}

@@

\subsection*{Our measurements are consistent with existing knowledge about HA's evolution and function}
How do the HA amino-acid preferences measured in our experiments relate to the evolution of H3N2 influenza virus in nature?
This question can be addressed by evaluating how well an experimentally informed codon substitution model (ExpCM) using our measurements describes H3N2 evolution compared to standard substitution models~\cite{bloom2017identification,hilton2017phydms}.
Table~\ref{tab:phydms} shows that the phylogenetic fit of the ExpCM using the across-replicate average of our measurements greatly outperforms conventional substitution models.
This result indicates that our experiments authentically capture some of the constraints on HA evolution.
A substitution model in which the amino-acid preferences were averaged across all sites (ExpCM, site avg.) performs no better than conventional substitution models, demonstrating the importance of site-specificity for the preferences.
The relative rate of nonsynonymous to synonymous substitutions (dN/dS or $\omega$) is $\ll$1 for conventional substitution models (Table~\ref{tab:phydms}), illustrating that these models need to account for constraints on natural HA evolution.
However, the relative rate of nonsynonymous to synonymous substitutions after accounting for the amino-acid preferences measured in our experiments ($\omega$ for the ExpCM) is close to one (Table~\ref{tab:phydms}), indicating that most purifying selection against nonsynonymous substitutions is accounted for in the deep mutational scanning.
The ExpCM stringency parameter~\cite{hilton2017phydms} is 2.47 (Table~\ref{tab:phydms}), indicating that natural selection favors the same amino acids as the experiments but with greater stringency.
Throughout the rest of this paper, we use the amino-acid preferences re-scaled~\cite{bloom2017identification,hilton2017phydms} by this stringency parameter.
These re-scaled preferences are shown in Figure~\ref{fig:logoplot}.
Furthermore, Figure~\ref{suppfig:diffprefs_logoplot} shows that when we compare these preferences to amino-acid frequencies from natural human H3N2 HA sequences, at most sites, there is little difference between the amino acids that are preferred in our experiments to their frequencies in nature.

@@

After annotating mutations and their frequencies on the phylogeny in this way (Figure~\ref{fig:frequency_trajectory}), it is visually obvious that there are relatively few circulating mutations that we measure to be strongly deleterious---and that such deleterious mutations rarely reach high frequency when they do occur.

We next sought to quantify the correlation between a mutation's experimentally measured effect and the maximum frequency it attained during natural evolution.
To calculate a given mutation's effect, we simply took the logarithm of the preference for the new amino-acid identity divided by the preference for the original amino-acid identity.
To minimize effects related to the genetic background of the strain used in the experiment, we excluded mutations closely related to the experimental strain itself and partitioned the remaining mutations into 1,022 mutations pre-dating and 299 mutations post-dating the Perth/2009 strain (Figure~\ref{suppfig:tree}).
We additionally excluded mutations from the post-Perth partition that were sampled in 2014 or after that have not had enough time for their evolutionary fates to be fully resolved.
We used these pre-Perth and post-Perth partitions to test the utility of our measurements for both post-hoc and prospective analyses, respectively.
We quantified the relationship between mutational effects and maximum mutation frequencies in the H3N2 phylogeny via Spearman rank correlation (Figure~\ref{fig:muteffect_maxfreq}A).
In both pre-Perth and post-Perth time periods, we found a modest, but statistically significant relationship between mutational effect and maximum mutation frequency (pre-Perth $\rho = 0.17$, post-Perth $\rho = 0.14$).

To restrict potential biases of our experimental measurements towards mutations from cell- or egg-passaged isolates, we evaluated mutations from unpassaged isolates.
Due to the limited availability of primary isolate sequences until the last decade, we analyzed mutations of unpassaged isolates from the post-Perth partition of the phylogenetic tree~\cite{mcwhite2016sequence}.
The correlation between mutational effect and maximum frequency was stronger for mutations from unpassaged post-Perth isolates ($\rho = 0.24$).
The similar effect sizes for both partitions shows that our experimental measurements can help explain the evolutionary fates of mutations in strains that post-date the experimentally studied strain, as well as to retrospectively analyze mutations that precede the experimental strain.

@@

It is important to emphasize that measurements of viral growth in cell culture do \emph{not} represent true fitness in nature.
Indeed, a vast amount of work in virology has chronicled the ways in which experiments can select for lab artifacts or fail to capture important pressures that are relevant in nature~\cite{daniels1985fusion,sun2010modifications,lee2013comparison,wu2017structural}.
As an example of this, although we identified G78D as a lab adaptation mutation and the wildtype Gly is not highly preferred at this site, this mutation never reaches above 20\% frequency in nature.
Mutations in viral genes other than HA are also important in determining strain success~\cite{memoli2009recent,raghwani2017selection}.
Given these caveats, it might seem surprising that measuring viral growth in cell culture can be informative about the success of viral strains in nature.
Yet, prior to our work, there were no comprehensive studies of the functional effects of mutations to H3 HA on any property that even resembled viral fitness in nature, and modeling work has either omitted the non-antigenic effects of mutations~\cite{sun2013using,harvey2016identification,neher2016prediction} or assumed that all non-epitope mutations had equivalent deleterious effects~\cite{luksza2014predictive}.
The strength of our measurements are not that they perfectly capture fitness in nature, but that they are systematic and quantitative---and so represent an improvement over no information at all.
We suspect that performing similar experiments using more realistic and complex selections (e.g., ferrets or primary human airway cultures) might further improve their utility and possibly their generalizability to more divergent strains.

@@

\subsection*{Quantification of Shannon entropy and relative solvent accessibility}
We calculated Shannon entropy, $h_r$ for site $r$ as $h_r = - \sum_{x} \pi_{r,x} \log(\pi_{r,x})$, where $\pi_{r,x}$ is the preference for amino acid $x$ at site $r$.

We quantified the absolute solvent accessibility of each site of the H3 HA (PDB 4O5N;~\cite{lee2014receptor}) or the H1 HA (PDB 1RVX;~\cite{gamblin2004structure}) structure using DSSP~\cite{kabsch1983}.
We then normalized to a relative solvent accessibility using the absolute accessibilities in~\cite{tien2013}.
